---
title: "Course Project"
author: "Veronica Vaca"
date: "31 de enero de 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The goal of this project is to predict the manner in which some people perform barbell lifts. We will use some variables in the data to do it. We are going to show:

- How the model was built.
- How was used cross validation in it.
- Expected out of sample error. 
- Conclusion and explanation. 
- Predict 20 different test cases.

## Data

The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har.

```{r,echo=FALSE}
#Load libraries
library(caret);library(rpart);library(rpart.plot);library(RColorBrewer)
library(rattle);library(randomForest);library(knitr);require(RANN)

#Download the data
if(!file.exists("pml-training.csv")){download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "pml-training.csv")}

if(!file.exists("pml-testing.csv")){download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv")}


#Read the training data and replace empty values by NA
trainingDataSet<- read.csv("pml-training.csv", sep=",", header=TRUE, na.strings = c("NA","",'#DIV/0!'))
testingDataSet<- read.csv("pml-testing.csv", sep=",", header=TRUE, na.strings = c("NA","",'#DIV/0!'))

```

## Cleaning the data

Removing variables that are not usable that have nearly zero variance, variables that are almost always NA, and variables that don’t make intuitive sense for prediction.

```{r, echo =F}

trainingDataSet <- trainingDataSet[,(colSums(is.na(trainingDataSet)) == 0)]

testingDataSet <- testingDataSet[,(colSums(is.na(testingDataSet)) == 0)]


numericalsIdx <- which(lapply(trainingDataSet, class) %in% "numeric")

preprocessModel <-preProcess(trainingDataSet[,numericalsIdx],method=c('knnImpute', 'center', 'scale'))
pre_trainingDataSet <- predict(preprocessModel, trainingDataSet[,numericalsIdx])
pre_trainingDataSet$classe <- trainingDataSet$classe

pre_testingDataSet <-predict(preprocessModel,testingDataSet[,numericalsIdx])


nzv <- nearZeroVar(pre_trainingDataSet,saveMetrics=TRUE)
pre_trainingDataSet <- pre_trainingDataSet[,nzv$nzv==FALSE]

nzv <- nearZeroVar(pre_testingDataSet,saveMetrics=TRUE)
pre_testingDataSet <- pre_testingDataSet[,nzv$nzv==FALSE]

```


## Building the model

For reproducibility we are going to set a seed and then subseting for having a training and a test set.

```{r}
set.seed(2000)

idxTrain<- createDataPartition(pre_trainingDataSet$classe, p=3/4, list=FALSE)
training<- pre_trainingDataSet[idxTrain, ]
validation <- pre_trainingDataSet[-idxTrain, ]
dim(training) ; dim(validation)

```

We will start with the Decision Trees

```{r}
modFitdt<-rpart(classe ~ ., data=training, method="class")
fancyRpartPlot(modFitdt,cex=.3,under.cex=0.2,shadow.offset=0)

```

Now we'll see the accuracy

```{r}
predictiondt <- predict(modFitdt, validation, type = c("raw"))

confusionMatrix(validation$classe,predictiondt)

```

We can see that Decission Trees are not a good model
Then we'll test the Random Forest

```{r}

modFitrf <- train(classe ~., method="rf", data=training, trControl=trainControl(method='cv'), number=5, allowParallel=TRUE, importance=TRUE )
modFitrf


```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
